Fishing for Spam

### 2024-06-15 code/eda-code/eda.qmd
Loading, cleaning, and exploring the datasets.

Density plots for the continuous outcomes -- email body length first.

@@

Now subject line length.

```{r}
p2 <- mydata %>% ggplot(aes(x=subjectlength, fill = label)) +
      geom_density(position = "dodge", alpha = 0.5) +
      xlim(0,150) + xlab("Length") + ylab("Density") +
      labs(title = "Subject Line Distribution", caption = "x-axis limited to a maximum of 150", fill = "Spam") +
      theme_tufte() +
      theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom")
plot(p2)
figure_file = here::here("results", "figures","subjectlength-distribution.png")
ggsave(filename = figure_file, plot=p2) 
```

Now the most common sender domains for spam emails.

```{r}
p3 <- mydata[which(mydata$label == "Yes" & !(mydata$senderdomain == "")),] %>%
      dplyr::count(senderdomain) %>%
      arrange(desc(n)) %>%
      head(25) %>%
      ggplot(aes(x = n, y = fct_reorder(senderdomain, n))) +
      geom_col(fill = "turquoise3", colour = "black", alpha = 0.65) +
      labs(title = "Spam Emails by Sender Domain") +
      ylab("") + xlab("Count") +
      theme_tufte() +
      theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "none")
plot(p3)
figure_file = here::here("results", "figures", "spam-by-domain.png")
ggsave(filename = figure_file, plot=p3) 
```
```{r}
p3 <- mydata[which(mydata$label == "No" & !(mydata$senderdomain == "")),] %>%
      dplyr::count(senderdomain) %>%
      arrange(desc(n)) %>%
      head(25) %>%
      ggplot(aes(x = n, y = fct_reorder(senderdomain, n))) +
      geom_col(fill = "red3", colour = "black", alpha = 0.35) +
      labs(title = "Legit Emails by Sender Domain") +
      ylab("") + xlab("Count") +
      theme_tufte() +
      theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "none")
plot(p3)
figure_file = here::here("results", "figures", "legit-by-domain.png")
ggsave(filename = figure_file, plot=p3) 
```

>>>>

Density plots for the continuous outcomes -- email body length first. We'll utilize the `ggplot2` visualization package, which offers a `geom_density` shape that is well suited to visualizing the distributions of multiple groups in the same plot. In our case that will be both spam and legitimate emails.

@@

And we'll produce the same visualization for subject line length.

```{r}
p2 <- mydata %>% ggplot(aes(x=subjectlength, fill = label)) +
      geom_density(position = "dodge", alpha = 0.5) +
      xlim(0,150) + xlab("Length") + ylab("Density") +
      labs(title = "Subject Line Distribution", caption = "x-axis limited to a maximum of 150", fill = "Spam") +
      theme_tufte() +
      theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom")
plot(p2)
figure_file = here::here("results", "figures","subjectlength-distribution.png")
ggsave(filename = figure_file, plot=p2) 
```
Now the most common sender domains for spam emails. We'll continue to utilize `ggplot2`, this time with the `geom_col` shape in order to produce a bar chart of counts.

```{r}
p3 <- mydata[which(mydata$label == "Yes" & !(mydata$senderdomain == "")),] %>%
      dplyr::count(senderdomain) %>%
      arrange(desc(n)) %>%
      head(25) %>%
      ggplot(aes(x = n, y = fct_reorder(senderdomain, n))) +
      geom_col(fill = "turquoise3", colour = "black", alpha = 0.65) +
      labs(title = "Spam Emails by Sender Domain") +
      ylab("") + xlab("Count") +
      theme_tufte() +
      theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "none")
plot(p3)
figure_file = here::here("results", "figures", "spam-by-domain.png")
ggsave(filename = figure_file, plot=p3) 
```
And similarly we'll produce a bar chart for legitimate emails.

```{r}
p3 <- mydata[which(mydata$label == "No" & !(mydata$senderdomain == "")),] %>%
      dplyr::count(senderdomain) %>%
      arrange(desc(n)) %>%
      head(25) %>%
      ggplot(aes(x = n, y = fct_reorder(senderdomain, n))) +
      geom_col(fill = "red3", colour = "black", alpha = 0.35) +
      labs(title = "Legit Emails by Sender Domain") +
      ylab("") + xlab("Count") +
      theme_tufte() +
      theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5), legend.position = "none")
plot(p3)
figure_file = here::here("results", "figures", "legit-by-domain.png")
ggsave(filename = figure_file, plot=p3) 
```

### 2024-06-15 products/manuscript/Manuscript.qmd
Loading, cleaning, and exploring the datasets.

By the end of our data loading and processing we arrived at a dataset that contains 7 character columns, 2 numeric columns, 2 factor columns, and a single boolean column -- taking care to ensure that our steps prevented generating any missing values where there hadn't previously been any. Ultimately this makes for a dataset with 48,295 rows across 11 features and our variable of interest. Complete steps and code can be located in the `.../code/processingprocessingcode.R` file found within the project.

@@

>>>>

By the end of our data loading and processing we arrived at a dataset that contains 7 character columns, 2 numeric columns, 2 factor columns, and a single boolean column -- taking care to ensure that our steps prevented generating any missing values where there hadn't previously been any. Ultimately this makes for a dataset with 48,295 rows across 11 features and our variable of interest. Complete steps and code can be located in the `.../code/processing/processingcode.R` file found within the project.

@@

A complete set of exploratory visualization steps utilizing `ggplot2` can be located in the `.../code/eda-code/eda.qmd` file found within the project.
